---
title: "EDA - Marium"
author: "Marium Tapal"
date: "3/13/2021"
output: html_document
---

```{r packages, include=FALSE}
library(tidyverse)
library(lubridate)
library(parcats)
library(ggforce)
library(patchwork)
```

```{r load-data, message=FALSE}
comm_data_Fri <- read_csv("~/Documents/Spring 2021/SDS 235/sds235-dc2/data/Communication Data/comm-data-Fri.csv")
comm_data_Sat <- read_csv("~/Documents/Spring 2021/SDS 235/sds235-dc2/data/Communication Data/comm-data-Sat.csv")
comm_data_Sun <- read_csv("~/Documents/Spring 2021/SDS 235/sds235-dc2/data/Communication Data/comm-data-Sun.csv")
```

```{r data-wrangling, message=FALSE}
glimpse(comm_data_Fri)

# change timestamp to date and times
# could have done this after rbind but if we want to use dfs separately 
comm_data_Fri$Timestamp <- parse_date_time(comm_data_Fri$Timestamp, "YmdHMS")
comm_data_Sat$Timestamp <- parse_date_time(comm_data_Sat$Timestamp, "YmdHMS")
comm_data_Sun$Timestamp <- parse_date_time(comm_data_Sun$Timestamp, "YmdHMS")

# add day column to data
comm_data_Fri <- comm_data_Fri %>% mutate(day = "Friday")
comm_data_Sat <- comm_data_Sat %>% mutate(day = "Saturday")
comm_data_Sun <- comm_data_Sun %>% mutate(day = "Sunday")

# Make one df
data <- rbind(comm_data_Fri, comm_data_Sat, comm_data_Sun)

# factor ids as levels instead of numeric
data <- data %>% mutate(to = as.factor(to),
                        from = as.factor(from))

# counts of data
sent_data <- data %>% group_by(from, day) %>% count() %>% filter(n>5000)
p1 <- ggplot(sent_data, aes(x=from, y = n, color = day)) + geom_point()
received_data <- data %>% group_by(to, day) %>% count() %>% filter(n>5000)
p2 <- ggplot(received_data, aes(x = to, y = n, color = day)) + geom_point()
# id numbers 839736 and 1278894 and external

p1 + p2

# data sent and received tog
data_from_to <- data %>% group_by(from, to, day) %>% count()
p3 <- ggplot(data_from_to, aes(x = to, y = n, color = day)) + geom_point()
```

In my exploratory data analysis above, the first things I figured out were that the id numbers 839736 and 1278894 and external person(s) had an excessive amount of messages sent and received. See plots p1 and p2.

These two id's don't match with the movement data - which suggests that these might be park service employees. (Hypothesis)

I further investigated this spike of messages and found out (in p3) that something was happening on Sunday, which might explain the excessive number of messages from and to the employees. 

Looking at the time line of these messages sent:

```{r}
data2 <- data %>% mutate(hour = hour(Timestamp),
                         minute = minute(Timestamp))
# saturday
hour_sat <- data2 %>% 
  filter(from == 839736 | from == 1278894) %>% 
  filter(day == "Saturday") %>% 
  group_by(from, hour) %>% 
  count()
p4 <- ggplot(hour_sat, aes(x = hour, y = n)) + geom_line()
p4 # pretty even

# sunday
hour_sun <- data2 %>% 
  filter(from == 839736 | from == 1278894) %>% 
  filter(day == "Sunday") %>% 
  group_by(from, hour) %>% 
  count()
p5 <- ggplot(hour_sun, aes(x = hour, y = n)) + geom_line()
p5 # peak looks like around noon and every two hours
ggsave("hours.jpg")

# look into minutes around on Sunday at noon
hourmin_sun <- data2 %>% 
  filter(from == 839736 | from == 1278894) %>% 
  filter(day == "Sunday", hour == 12) %>% 
  group_by(from, minute) %>% 
  count()
p6 <- ggplot(hourmin_sun, aes(x = minute, y = n)) + geom_line()
p6 
# between the first 30 mins - 12-12:30?
# the spikes in every hour could be regular messages sent by employees?
ggsave("noon.jpg")


# check with received messages
# saturday
hour_sat_to <- data2 %>% 
  filter(to == 839736 | to == 1278894) %>% 
  filter(day == "Saturday") %>% 
  group_by(to, hour) %>% 
  count()
p7 <- ggplot(hour_sat_to, aes(x = hour, y = n)) + geom_line()
p7 # pretty even

# sunday
hour_sun_to <- data2 %>% 
  filter(to == 839736 | to == 1278894) %>% 
  filter(day == "Sunday") %>% 
  group_by(from, hour) %>% 
  count()
p8 <- ggplot(hour_sun_to, aes(x = hour, y = n)) + geom_line()
p8 # peak sunday afternoon too

# look into minutes around on Sunday at 11 am - 4 pm
# manually changed hour from 11 to 16 to see the changes
hourmin_sun_to <- data2 %>% 
  filter(to == 839736 | from == 1278894) %>% 
  filter(day == "Sunday", hour == 12) %>% 
  group_by(from, minute) %>% 
  count()
p9 <- ggplot(hourmin_sun_to, aes(x = minute, y = n)) + geom_line()
p9 
# between the first 30 mins - 12:00-15:40?
# the spikes in every hour could be regular messages sent by employees?
```


```{r}
# # look into id numbers 839736 and 1278894 and external more
# sent_ids <- c(839736, 1278894)
# received_ids <- c(839736, 1278894, "external")
# 
# location <- data %>% filter(to == c(839736, 1278894))
# ggplot(location, aes(x = location, fill = to)) + geom_bar(position = "dodge")
```


- look into hours on sunday

ideas:
- communication by location - network graph?
- group by ids? how many people were there
